---
title: "My answers"
author: "My name"
date: "2021-12-02"
output: html_document
---




# 1. Simple linear model

> In order to tryout linear models in _R_ we are going to use the `blood_fat` dataset which contains the *age* (in years), *weight* (in kg) and  measured *fat concentrations* (units were not mentioned) in blood samples of different subjects.

## Relation of blood fat content and age

We would like to determine whether the blood concentration in fat is related to the age of the subjects.

(@) load the data (`csv`) available [here](data/blood_fat.csv) in R

```{r}
# Write your answer here
```



### Visualization

(@) If the goal would be to guess the fat concentration knowing someones age, find out which are the **response** and **predictor** variables and draw the according scatter plot.
(@) Add a regression line (without the confidence interval ribbon).
```{r}
# Write your answer here
```


(@) Add a new column to the `blood_fat` data frame containing the expected fat levels from the linear model.
    - For each subject in the data frame, add these predicted values as **red points** on your previous plot.
    - Using `geom_segment()`, connect the expected values to the measured values as **dotted lines**.

```{block, opts.label = "tip"}
remember to use the appropriate `broom` function after the linear model, to fetch all those information in one tibble
```


```{r}
# Write your answer here
```


(@) Calculate the **slope** and **intercept** of the regression line

```{r}
# Write your answer here
```


(@) draw a **dashed lightblue** line using **explicitly** the values you calculated in the previous question

```{r}
# Write your answer here
```



### Calculate $R^2$

You learned that $R^2$ can be calculated as follows:

  $$R^2 = 1- \frac{\sum(y_i - \hat{y_i})^2}{\sum(y_i - \bar{y})^2} = 1- \frac{RSS}{TSS}$$

(@) The length of the dotted lines you just represented are related to a term of this equation. Which one?
(@) Using `mutate()`, add the length of each dotted line to the `blood_fat` data frame.

```{r}
# Write your answer here
```


(@) What do these length represent? Which functions in _R_ generates these values?

```{r}
# Write your answer here
```



(@) Draw a **darkgreen** horizontal line showing the mean fat concentration of all measures.
(@) Use `geom_segment()` to connect all points representing the real measures to their projection on this horizontal line (using a **darkgreen** color with **alpha = 0.2** and a **size of 2**).

```{r}
# Write your answer here
```


(@) Using `mutate()`, add the length of each green translucid line to the `blood_fat` data frame. What do they represent? And the sum of their squared length?

```{block, opts.label = "tip"}
Those lines are deviation to the global mean of the response
```

```{r}
# Write your answer here
```


(@) Calculate $RSS$, $TSS$ and $R^2$

```{r}
# Write your answer here
```


```{r}
# Write your answer here
```


(@) Is there really a relationship between blood fat content and age?
(@) What does the value of $R^2$ tell you?

```{r}
# Write your answer here
```


### Checking the residuals of the model

(@) Residuals's mean
    + What is the expectation for the residuals's mean?
    + Compute the residuals's mean for the fat explained by age model.
```{r}
# Write your answer here
```


(@) Can the measures appropriately be modelled in this way? Draw two diagnosis plots using `ggplot2`
    + In the first draw the residuals on the _y_ axis and the estimated values on the _x_ axis
    + Your second one should be a quantile-quantile plot.


```{block, opts.label = "tip"}
The package `ggfortify` and the function `autoplot(fit)` can produce the classic 4 diagnostic plots with no efforts
```

```{r}
# Write your answer here
```


### Relation of blood fat content and weight

(@) Change predictor and use `weight` instead of `age` to predict the fat concentration

```{r}
# Write your answer here
```


(@) Check the summary and diagnostic plots for this regression

```{r}
# Write your answer here
```


```{r}
# Write your answer here
```


(@) Can the blood content in fat be explained by the weight of the subject?

```{r}
# Write your answer here
```



# 2. Multiple linear regression 
## Simulated data

We are going to manipulate data where the relationship between the variables is simulated. This example comes from Gareth _et al._ 2013.

### Generate the data

(@) Use the following code to generate the data:

```{r generate}
set.seed(1)
simul1 <- tibble(x1 = runif(100),
                     x2 = 0.5 * x1 + rnorm(100) / 10,
                     y  = 2 + 2 * x1 + 0.3 * x2 +(rnorm(100)))
```

### Variable _y_ is using a linear model

(@) Write the equation (as text in your markdown document)
(@) Write the regression coefficients (as text in your markdown document)
    
```{r}
# Write your answer here
```


(@) Plot the relationship between _x1_ and _x2_. You might want to display all relationships using `ggpairs`

```{block, opts.label = "tip"}
`ggpairs` comes from the package `GGally`
```


```{r}
# Write your answer here
```


(@) What is the Pearson's correlation coefficient between _x1_ and _x2_?

```{r}
# Write your answer here
```


(@) Will the correlation between predictors be a problem for the linear regression you just wrote down?
(@) Name this effect.

```{r}
# Write your answer here
```


(@) Fit a linear model explaining _y_ by _x1 + x2_ and describe the summary results.\
(@) Do you accept $H_0: \hat{\beta_1} = 0$ and  $H_0: \hat{\beta_2} = 0$?

```{r}
# Write your answer here
```


(@) Fit a linear model explaining _y_ by _x1_ alone.
(@) Do you accept $H_0: \hat{\beta_1} = 0$?
What happened?

```{r}
# Write your answer here
```


(@) Fit a linear model explaining _y_ by _x2_ alone.\
(@) Do you accept $H_0: \hat{\beta_1} = 0$?\
What happened?

```{r}
# Write your answer here
```


### One observation was unfortunately forgotten and is:

`x1 = 0.1, x2 = 0.8, y = 6`
(@) Add it to the existing `simul1` and store the result as `simul2`.

```{block, opts.label = "tip"}
Have a look at the function `add_row()` defined in the package `tibble`
```



```{r}
# Write your answer here
```


(@) Plot the relationship between _x1_ and _x2_ again using the updated dataset `simul2`.

```{r}
# Write your answer here
```


(@) Fit the previous linear model again using `simul2`

```{r}
# Write your answer here
```


```{r}
# Write your answer here
```

